PANCHANGE_data_preprocessing

### 2020-08-13 PANCHANGE_data_preprocessing2020-07-15.Rmd
First stage of adapting MHD section for NBR. Reading in data and cleaning depressive disorders chunks completed.

@@

```{r glad mhd depressive_disorders}
mhd.raw.id <- mhd.raw.id %>%
  mutate(
    depressive_disorders_numeric =
      case_when(
        (mhd.depression_numeric == "0" &
        (mhd.pregnancy_depression_numeric == "0" | is.na(mhd.pregnancy_depression_numeric)) & 
        (mhd.premenstrual_dysphoric_disorder_pmdd_numeric == "0" | is.na(mhd.premenstrual_dysphoric_disorder_pmdd_numeric))) ~ 0,
        mhd.depression_numeric == "1" ~ 1,
        mhd.pregnancy_depression_numeric == "1" ~ 1,
        mhd.premenstrual_dysphoric_disorder_pmdd_numeric == "1" ~ 1))

@@

Anxiety disorders group
```{r glad mhd anxiety_disorders}
mhd.raw.id <- mhd.raw.id %>%
  mutate(
    anxiety_disorders_numeric =
      case_when(
        mhd.anxiety_nerves_or_generalised_anxiety_disorder_numeric == "1" ~ 1,
        mhd.social_anxiety_or_social_phobia_numeric == "1" ~ 1,
        mhd.specific_phobia_e.g._phobia_of_flying_numeric == "1" ~ 1,
        mhd.agoraphobia_numeric == "1" ~ 1,
        mhd.panic_disorder_numeric == "1" ~ 1,
        mhd.panic_attacks_numeric == "1" ~ 1,
        (mhd.anxiety_nerves_or_generalised_anxiety_disorder_numeric == "0" &
        mhd.social_anxiety_or_social_phobia_numeric == "0" &
        mhd.specific_phobia_e.g._phobia_of_flying_numeric == "0" &
        mhd.agoraphobia_numeric == "0" &
        mhd.panic_disorder_numeric == "0" &
        mhd.panic_attacks_numeric == "0") ~ 0
      )
  )

@@

mhd.raw.id %>%
  freq(anxiety_disorders,
       cumul = F)
```

>>>>

## EDGI
```{r edgi mhd cols}
if(EDGI == TRUE) {
edgi.mhd.raw <- readRDS(file = paste0(data.raw_path,"/edgi/mhd_edgi.rds"))
dim(edgi.mhd.raw)
colnames(edgi.mhd.raw)
}
```

```{r edgi mhd cleaning}
if(EDGI == TRUE) {
edgi.mhd.raw.id <- edgi.mhd.raw %>%
  drop_na(externalDataReference) %>% # Drop NAs
  distinct(externalDataReference, .keep_all = TRUE) %>% # Remove duplicates based on ID
  separate(externalDataReference, into = c("Sample", "ID"), sep = 4) %>% # Split ID in Sample and Number
  mutate(ID = as.numeric(ID),
         startDate.prepandemic = startDate) %>% # Rename startDate
  select(
    Sample, # Sample
    ID, # ID
    starts_with("mhd.")
  ) %>%
  add_numeric(., exclude = exclude_cols) %>%
  mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
  mutate_if(is.numeric, ~na_if(., -99)) %>%
  mutate_if(is.numeric, ~na_if(., -77)) %>%
  mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
  mutate_if(is.factor, ~na_if(., "Don't know")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

# Inspect dimensions
dim(edgi.mhd.raw.id)
# Inspect colnames
colnames(edgi.mhd.raw.id)
#Differences
dim(edgi.mhd.raw)[1]-dim(edgi.mhd.raw.id)[1]
}
```

```{r edgi mhd copy}
if(EDGI == TRUE) {
  mhd.raw.id <- edgi.mhd.raw.id
}
```

## NBR
```{r nbr mhd cols}
if(NBR == TRUE) {
nbr.mhd.raw <- readRDS(file = paste0(data.raw_path,"/nbr/mhd_nbr.rds"))
dim(nbr.mhd.raw)
colnames(nbr.mhd.raw)
}
```

```{r nbr mhd cleaning}
if(NBR == TRUE) {
exclude_cols <- c("Sample",
                  "ID")

nbr.mhd.raw.id <- nbr.mhd.raw %>%
  drop_na(subjectid) %>% # Drop NAs
  distinct(subjectid, .keep_all = TRUE) %>% # Remove duplicates based on ID
  separate(subjectid, into = c("Sample", "ID"), sep = 4) %>% # Split ID in Sample and Number
  mutate(ID = as.numeric(ID),
         startDate.prepandemic = startDate) %>% # Rename startDate
  select(
    Sample, # Sample
    ID, # ID
    starts_with("mhd.")
  ) %>%
  add_numeric(., exclude = exclude_cols) %>%
  mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
  mutate_if(is.numeric, ~na_if(., -99)) %>%
  mutate_if(is.numeric, ~na_if(., -77)) %>%
  mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
  mutate_if(is.factor, ~na_if(., "Don't know")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

# Inspect dimensions
dim(nbr.mhd.raw.id)
# Inspect colnames
colnames(nbr.mhd.raw.id)
#Differences
dim(nbr.mhd.raw)[1]-dim(nbr.mhd.raw.id)[1]
}
```

```{r nbr mhd copy}
if(NBR == TRUE) {
  mhd.raw.id <- nbr.mhd.raw.id
}
```

@@

```{r mhd depressive_disorders}
## CHECK THAT THIS CHUNK WORKS FOR EDGI AS WELL (confirmed that it works for NBR)
if(GLAD = TRUE | EDGI = TRUE | NBR = TRUE)
mhd.raw.id <- mhd.raw.id %>%
  mutate(
    depressive_disorders_numeric =
      case_when(
        (mhd.depression_numeric == "0" &
        (mhd.pregnancy_depression_numeric == "0" | is.na(mhd.pregnancy_depression_numeric)) & 
        (mhd.premenstrual_dysphoric_disorder_pmdd_numeric == "0" | is.na(mhd.premenstrual_dysphoric_disorder_pmdd_numeric))) ~ 0,
        mhd.depression_numeric == "1" ~ 1,
        mhd.pregnancy_depression_numeric == "1" ~ 1,
        mhd.premenstrual_dysphoric_disorder_pmdd_numeric == "1" ~ 1))

@@

Anxiety disorders group
```{r glad mhd anxiety_disorders}
if(GLAD = TRUE) {
mhd.raw.id <- mhd.raw.id %>%
  mutate(
    anxiety_disorders_numeric =
      case_when(
        mhd.anxiety_nerves_or_generalised_anxiety_disorder_numeric == "1" ~ 1,
        mhd.social_anxiety_or_social_phobia_numeric == "1" ~ 1,
        mhd.specific_phobia_e.g._phobia_of_flying_numeric == "1" ~ 1,
        mhd.agoraphobia_numeric == "1" ~ 1,
        mhd.panic_disorder_numeric == "1" ~ 1,
        mhd.panic_attacks_numeric == "1" ~ 1,
        (mhd.anxiety_nerves_or_generalised_anxiety_disorder_numeric == "0" &
        mhd.social_anxiety_or_social_phobia_numeric == "0" &
        mhd.specific_phobia_e.g._phobia_of_flying_numeric == "0" &
        mhd.agoraphobia_numeric == "0" &
        mhd.panic_disorder_numeric == "0" &
        mhd.panic_attacks_numeric == "0") ~ 0
      )
  )

@@

mhd.raw.id %>%
  freq(anxiety_disorders,
       cumul = F)
}
```

```{r nbr mhd anxiety_disorders}
if(NBR = TRUE) {
mhd.raw.id <- mhd.raw.id %>%
  mutate(
    anxiety_disorders_numeric =
      case_when(
        mhd.anxiety_nerves_or_generalised_anxiety_disorder_numeric == "1" ~ 1,
        mhd.social_anxiety_or_social_phobia_numeric == "1" ~ 1,
        mhd.specific_phobia_e.g._phobia_of_flying_numeric == "1" ~ 1,
        mhd.agoraphobia_numeric == "1" ~ 1,
        mhd.panic_disorder_numeric == "1" ~ 1,
        mhd.panic_attacks_numeric == "1" ~ 1,
        (mhd.anxiety_nerves_or_generalised_anxiety_disorder_numeric == "0" &
        mhd.social_anxiety_or_social_phobia_numeric == "0" &
        mhd.specific_phobia_e.g._phobia_of_flying_numeric == "0" &
        mhd.agoraphobia_numeric == "0" &
        mhd.panic_disorder_numeric == "0" &
        mhd.panic_attacks_numeric == "0") ~ 0
      )
  )

mhd.raw.id %>%
  freq(anxiety_disorders,
       cumul = F)
}
```
